Title

### 2013-04-02 inc/model_results/shapes/shapes.tex
more tweaks to shapes section

The initial `space-time' eccentricity (defined in \ref{eq:define_xi}), $\xi ( 0 ) = v_{\smallzero} \tau / w $ , of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \sigma_{ \smallT }(0) = \sigma_{z} ( 0 ) $), oblate or `disk-like' ($ \sigma_{ \smallT } ( 0 ) > \sigma_{ z } ( 0 ) $), and prolate or `cigar-like' ($ \sigma_{ \smallT } ( 0 ) < \sigma_{ z } ( 0 ) $) pulses.
%TODO rewrite using w?
This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w_{\smallT}^2 w_z = $ 1mm$^3$.
Pulses used in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} have just one electron per pulse (i.e. $ N = 10^{0} $), while \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} have a much higher $ N = 10^{6} $ electrons in the bunch.
Since the initial pulse volume is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1}.
The pulse velocity is $ v_{\smallzero} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.

@@

The bunch propagation dynamics for negligible pulse charge density case shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
In particular, in the initially spherical case ($ \sigma_{\smallT} (0) = \sigma_{z} (0) $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp.
Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both oblate and prolate pulses are initially driven towards the spherical regime.
In all cases, this intrinsic electron pulse broadening due to non-zero $ \eta_{\alpha} $ generates a momentum chirp across the electron bunch; in the longitudinal dimension the faster electrons photoemitted with larger excess energy leading the slower electrons.

For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects can dominate the propagation dynamics.
\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
In the spherical case, space-charge induced bunch broadening begins to overwhelm the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing a more uniform three-dimensional `Coulomb explosion' of the electron pulse with the overall bunch broadening rate much greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation. This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
% the following sentance is no longer true, still perhaps should include something like it
%In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.

>>>>

The initial `space-time' eccentricity (defined in \ref{eq:define_xi}),
\begin{equation}
  \xi ( 0 ) = \sqrt{ \frac{ \sigma_{z}(0) }{ \sigma_{\smallT}(0) } } = \frac{ w_z(0) }{ w(0) } = \frac{ v_{\smallzero} \tau(0) }{ w(0) } \,\text{,}
\end{equation}
of the photo-generated electron pulse plays a major role in the bunch propagation dynamics, particularly when space-charge effects dominate.
\ref{fig:compare_shape} illustrates the typical dynamics of initially spherical ($ \xi(0) = 0 $), oblate or `disk-like' ($ \xi(0) < 1 $), and prolate or `cigar-like' ($ \xi(0) > 1 $) pulses.
%TODO rewrite using w?
This simulation uses the idealized pulse conditions described above, where the initial bunch dimensions are given by the ellipticities $ \xi(0) = \{ 1/3 , 1 , 3 \} $ and in each case the product of the initial HW1/eM bunch widths $w^2 w_z = $ 1mm$^3$.
Pulses used in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} have just one electron per pulse (i.e. $ N = 10^{0} $), while \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} have a much higher $ N = 10^{6} $ electrons in the bunch.
Since the initial pulse volume is consistent between all runs, initial peak pulse charge density is six orders of magnitude greater in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} than in \ref{fig:compare_shape}\subref{fig:compare_shape_N1}.
The pulse velocity is $ v_{\smallzero} \approx c/3 $, i.e. a DC photogun with $V = 30\text{kV} $.

@@

The bunch propagation dynamics for negligible pulse charge density case shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1} display the trends expected from the initial photoemission conditions.
In particular, in the initially spherical case ($ \xi(0) = 1 $), the transverse bunch width $ \sqrt{2 \sigma_{\smallT}} $ expands approximately twice as rapidly as the longitudinal length $ \sqrt{2 \sigma_{z}} $ due to the factor of two difference in the initial momentum variances $ \Delta p_{\alpha} $ and the artificially zero initial pulse chirp.
Naturally, for initially non-spherical electron bunches, the pulse propagation model also shows that the relative bunch broadening $ \sqrt{ \sigma_{\alpha} / \sigma_{\alpha} (0) } $ for comparable $ \eta_{\alpha} $ is greatest for the smallest initial $ \sigma_{\alpha} (0) $; that is both oblate and prolate pulses are initially driven towards the spherical regime.
In all cases, this intrinsic electron pulse broadening due to non-zero $ \eta_{\alpha} $ generates a momentum chirp across the electron bunch; in the longitudinal dimension the faster electrons photoemitted with larger excess energy leading the slower electrons.

For an electron pulse with a higher initial space-charge density, intra-pulse Coulomb effects may dominate the propagation dynamics.
\ref{fig:compare_shape}\subref{fig:compare_shape_N1e6}, shows the predicted time evolution dynamics of electron pulses with the same initial size and photoemission conditions as \ref{fig:compare_shape}\subref{fig:compare_shape_N1}, but with $ N = 10^{6} $ electrons --- the number required for single-shot diffraction pattern measurements in a DTEM.\cite{berger_dc_2009,armstrong_practical_2007}
The significant influence of space-charge effects for the higher initial peak pulse charge density is clearly evident.
In the spherical case, space-charge induced bunch broadening begins to overwhelm the initial difference between $\eta_{\smallT}$ and $\eta_{z}$, producing a more uniform three-dimensional `Coulomb explosion' of the electron pulse with the overall bunch broadening rate much greater than that for $ N = 10^{0} $ (\ref{fig:compare_shape}\subref{fig:compare_shape_N1}).
For $ \xi (0) > 1 $ (or $ \xi (0) < 1 $), the smaller initial transverse (longitudinal) bunch dimension experiences the greatest relative expansion in free-space propagation. This is to be expected, because the initial $\eta_{\alpha}$ are non-zero (but comparable in magnitude) while Coulomb effects will produce differential pulse broadening in the dimension with the largest internal space-charge field (i.e., the dimension with the largest electrostatic potential gradient).
% the following sentance is no longer true, still perhaps should include something like it
%In fact, both the $ N = 10^{6} $ oblate and prolate pulse examples shown in \ref{fig:compare_shape}\subref{fig:compare_shape_N1e6} transition through a spherical bunch shape ($ \xi \approx 1 $) after a time of flight of $\sim$0.5ns, at which point the peak pulse charge density has already been reduced by a factor of $\sim$100.
The initial impulse induced by the strong space-charge field then provides for continued linear expansion (in time), resulting in a linear pulse chirp in the dimension associated with the smallest initial HW1/eM bunch width\cite{michalik_analytic_2006,berger_dc_2009} $ \sqrt{ 2 \sigma_{\alpha} (0) } $.
